---
title: "About"
---

```{r}
#| message: false

source("setup.R")

counts_type <- listings_summary |>
  group_by(inventory_type) |>
  summarise(n = sum(n))
n_new <- counts_type[counts_type$inventory_type == 'new', ]$n
n_used <- counts_type[counts_type$inventory_type == 'used', ]$n
n_new <- scales::comma(n_new)
n_used <- scales::comma(n_used)
```

This dashboard provides summary information about various trends in the U.S. automobile market. The primary data source is `r n_new` new and `r n_used` used vehicle listings from over 100,000 dealerships across all 50 states in the U.S. The original data are licensed from marketcheck.com and cannot be made public, but this website provides public access to various summaries of the original listings data.

The open-source project was created and is maintained by [Dr. John Paul Helveston](//https://jhelvy.com), an Associate Professor in the Engineering Management and Systems Engineering Department at The George Washington University.

## Vehicle Miles Traveled (VMT)

The [Daily VMT](vmt-daily.qmd) page shows the distribution of daily VMT across all vehicles in the dataset. Each chart displays quantiles of daily mileage for a given combination of powertrain and vehicle type. For example, the 0.5 quantile represents the median daily mileage (half of vehicles drive less than this amount and half drive more). 

The [VMT by Age](vmt-age.qmd) page shows median cumulative vehicle miles traveled (VMT) by vehicle age. Each chart displays the 50th quantile (median) of total mileage for a given combination of powertrain and vehicle type as vehicles age over time. 

For both of the VMT pages, we estimate VMT by dividing the odometer readings (mileage) of used vehicles by their age, which gives an approximation of the VMT over time. We estimate the age of the vehicle as the listing date minus the production date, which we estimate as June 30 of the year previous to the model year (e.g., we estimate a 2020 model year vehicle was produced June 30, 2019). We summarize this quantity by taking quantiles of the observations at each age bin (in 3-month intervals). For daily VMT, we use all vehicles regardless of age and compute the VMT for every quantile from 1 to 99. For VMT by age, we only include the 25th, 50th, and 75th quantile.  

## Depreciation

The [Depreciation](depreciation.qmd) page shows how vehicles retain value over time across different powertrains and vehicle types. Each chart displays the retention rate, which is the ratio of the current listing price to the original MSRP (Manufacturer's Suggested Retail Price). A retention rate of 1.0 means the vehicle retains 100% of its original value, while 0.5 means it has lost half its value.

We estimate the retention rate by dividing the listing price of used vehicles by the original MSRP for that make, model, and trim. Vehicle age is estimated as the listing date minus the production date, which we estimate as June 30 of the year previous to the model year (e.g., we estimate a 2020 model year vehicle was produced June 30, 2019). We summarize this quantity by taking the 25th, 50th, and 75th quantiles of the observations for each combination of powertrain and vehicle type at each age bin (in 3-month intervals). The charts display the 50th quantile (median) retention rate.

## Registrations

The [Registrations](registrations.qmd) page shows light-duty vehicle registration counts by state, powertrain, and year. Data are sourced from the [Alternative Fuels Data Center (AFDC)](https://afdc.energy.gov/vehicle-registration), which compiles registration data from IHS Markit. The page includes a choropleth map of the U.S. showing registration counts or market share for a selected powertrain and year, as well as a bar chart showing the trend over time for a selected state. Users can toggle between absolute registration counts and market share (the fraction of total registrations in each state attributed to a given powertrain). Clicking on a state in the map updates the bar chart to show that state's trend.

## Market Concentration

The [Market Concentration](market-concentration.qmd) page uses the [Herfindahl–Hirschman Index (HHI)](https://en.wikipedia.org/wiki/Herfindahl%E2%80%93Hirschman_Index) to measure how concentrated or competitive each powertrain segment is over time. The HHI ranges from 0 to 1, where values below 0.15 indicate a competitive market, values between 0.15 and 0.25 indicate moderate concentration, and values above 0.25 indicate a highly concentrated market. Concentration is measured across three dimensions: by make, by vehicle type, and by price bin. See the [HHI Explained](hhi-explained.qmd) page for an interactive tool that illustrates how HHI relates to market share distributions.

<!-- ## Listings

Detailed analysis of vehicle listings data, including pricing trends, availability, and inventory levels across different segments. Shows current market inventory patterns and pricing across vehicle types and powertrains.

## Dealerships Inventories

Explores dealership inventory patterns, stock levels, and geographic distribution of vehicle availability. Shows how dealer inventory is distributed across regions and vehicle types. -->

## Listing Counts

The table below summarises how many observations of vehicle listings were included in the summary data presented on this website.

```{r}
#| message: false

listings_summary |>
  mutate(
    inventory_type = tools::toTitleCase(inventory_type),
    count = scales::comma(n)
  ) |>
  select(-n) |>
  rename(
    Market = inventory_type,
    Powertrain = powertrain,
    `Vehicle Type` = vehicle_type
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 55,
    showPageSizeOptions = TRUE
  )
```