---
title: "About"
---

```{r}
#| message: false

library(dplyr)

data_url <- "https://raw.githubusercontent.com/vehicletrends/vehicletrends/refs/heads/main/data-raw/listings_summary.csv"
listing_summary <- read.csv(data_url)
counts_type <- listing_summary |>
  group_by(inventory_type) |>
  summarise(n = sum(n))
n_new <- counts_type[counts_type$inventory_type == 'new', ]$n
n_used <- counts_type[counts_type$inventory_type == 'used', ]$n
n_new <- scales::comma(n_new)
n_used <- scales::comma(n_used)
```

This dashboard provides summary information about various trends in the U.S. automobile market. The primary data source is `r n_new` new and `r n_used` used vehicle listings from over 100,000 dealerships across all 50 states in the U.S. The original data are licensed from marketcheck.com and cannot be made public, but this website provides public access to various summaries of the original listings data.

The open-source project was created and is maintained by [Dr. John Paul Helveston](https://jhelvy.com), an Associate Professor in the Engineering Management and Systems Engineering Department at The George Washington University.

## Vehicle Miles Traveled (VMT)

The [Daily VMT](daily-vmt.qmd) page shows the distribution of daily VMT across all vehicles in the dataset. Each chart displays quantiles of daily mileage for a given combination of powertrain and vehicle type. For example, the 0.5 quantile represents the median daily mileage (half of vehicles drive less than this amount and half drive more). 

The [VMT by Age](age-vmt.qmd) page shows median cumulative vehicle miles traveled (VMT) by vehicle age. Each chart displays the 50th quantile (median) of total mileage for a given combination of powertrain and vehicle type as vehicles age over time. 

For both of the VMT pages, we estimate VMT by dividing the odometer readings (mileage) of used vehicles by their age, which gives an approximation of the VMT over time. We estimate the age of the vehicle as the listing date minus the production date, which we estimate as June 30 of the year previous to the model year (e.g., we estimate a 2020 model year vehicle was produced June 30, 2019). We summarize this quantity by taking quantiles of the observations. For daily VMT, we use all vehicles regardless of age and compute the VMT for every quantile from 1 to 99. For VMT by age, we only include the 25th, 50th, and 75th quantile.  

## Depreciation

Examines how vehicles retain value over time across different powertrains. Shows the percentage of original price (MSRP) retained at each vehicle age, revealing which powertrains hold their value best and how quickly vehicles depreciate in their first 5 years.

## Market Concentration

Tracks market share and competitive dynamics across manufacturers within each powertrain category. Measures how concentrated or competitive each segment is, showing whether markets are dominated by a few manufacturers or have healthy competition across many brands.

## Listings

Detailed analysis of vehicle listings data, including pricing trends, availability, and inventory levels across different segments. Shows current market inventory patterns and pricing across vehicle types and powertrains.

## Dealerships Inventories

Explores dealership inventory patterns, stock levels, and geographic distribution of vehicle availability. Shows how dealer inventory is distributed across regions and vehicle types.

## Listing Counts

The table below summarises how many observations of vehicle listings were included in the summary data presented on this website.

```{r}
#| message: false

library(dplyr)

data_url <- "https://raw.githubusercontent.com/vehicletrends/vehicletrends/refs/heads/main/data-raw/listings_summary.csv"

read.csv(data_url) |>
  mutate(
    inventory_type = tools::toTitleCase(inventory_type),
    count = scales::comma(n)
  ) |>
  select(-n) |>
  rename(
    Market = inventory_type,
    Powertrain = powertrain,
    `Vehicle Type` = vehicle_type
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 55,
    showPageSizeOptions = TRUE
  )
```