---
title: "Cite"
---

If you use this dashboard or its data in your research or publications, please cite the relevant work below.

## Dashboard

> Helveston, John Paul (2026). *US Vehicle Trends Dashboard*. https://vehicletrends.us

```bibtex
@misc{helveston2026vehicletrends,
  author = {Helveston, John Paul},
  title  = {{US Vehicle Trends Dashboard}},
  year   = {2026},
  url    = {https://vehicletrends.us}
}
```

## R Data Package

The summary data powering this dashboard are also available as the [{vehicletrends}](https://pkg.vehicletrends.us/) R package. If you use the package data directly, please also cite it. You can get the citation information by typing `citation("vehicletrends")` into R:

```{r}
#| echo: true
#| eval: false

citation('vehicletrends')
```

```{r}
#| echo: true
#| eval: false

To cite vehicletrends in publications use:

  John Paul Helveston (2026). vehicletrends: Data on Vehicle Trends in
  the USA. R package version 0.0.3.
  https://github.com/vehicletrends/vehicletrends

A BibTeX entry for LaTeX users is

  @Manual{,
    title = {vehicletrends: Data on Vehicle Trends in the USA},
    author = {John Paul Helveston},
    year = {2026},
    note = {R package version 0.0.3},
    url = {https://github.com/vehicletrends/vehicletrends},
  }
```