---
title: "Market Concentration"
---

```{r}
#| label: setup
#| include: false

source("setup.R")

data <- vehicletrends::hhi |>
  filter(!(group_var == "price_bin" & group_level == "Other")) |>
  mutate(lower = pmax(lower, 0))

hhi_labels <- c(
  "make" = "Make",
  "vehicle_type" = "Vehicle Type",
  "price_bin" = "Price Bin"
)

make_boxplot <- function(data, gv, hv) {
  plot_data <- data |>
    filter(group_var == gv, hhi_var == hv)

  years <- sort(unique(plot_data$listing_year))
  levels <- sort(unique(plot_data$group_level))

  # Chart height based on number of categories
  chart_height <- paste0(max(300, length(levels) * 50 + 120), "px")

  # Left margin based on longest label
  grid_left <- max(100, max(nchar(levels)) * 7.5 + 30)

  # Build timeline options (one per year)
  opts <- lapply(years, function(yr) {
    yr_data <- plot_data |> filter(listing_year == yr)

    box_data <- lapply(levels, function(lvl) {
      row <- yr_data |> filter(group_level == lvl)
      if (nrow(row) == 0) return(list(0, 0, 0, 0, 0))
      c(row$lower[1], row$q25[1], row$median[1], row$q75[1], row$upper[1])
    })

    list(series = list(list(data = box_data)))
  })

  echarts4r::e_charts(height = chart_height) |>
    echarts4r::e_list(list(
      baseOption = list(
        timeline = list(
          data = as.character(years),
          autoPlay = FALSE,
          bottom = 0,
          playInterval = 1500
        ),
        tooltip = list(trigger = "item"),
        grid = list(left = grid_left, bottom = 80),
        yAxis = list(
          type = "category",
          data = levels,
          inverse = TRUE
        ),
        xAxis = list(
          type = "value",
          name = "HHI",
          min = 0
        ),
        series = list(
          list(type = "boxplot")
        ),
        toolbox = list(
          feature = list(saveAsImage = list())
        )
      ),
      options = opts
    ))
}
```

{{< include chunks/loading-notice.qmd >}}

The [Market Concentration](market-concentration.qmd) page uses the [Herfindahl–Hirschman Index (HHI)](https://en.wikipedia.org/wiki/Herfindahl%E2%80%93Hirschman_Index) to measure how concentrated or competitive each market segment is over time. The HHI ranges from 0 to 1, where values below 0.15 indicate a competitive market, values between 0.15 and 0.25 indicate moderate concentration, and values above 0.25 indicate a highly concentrated market. Concentration is measured across three dimensions: by make, by vehicle type, and by price bin. Each boxplot summarizes the distribution of HHI values across DMAs for a given year. Use the timeline slider below each chart to switch between years. See the [HHI Explained](hhi-explained.qmd) page for an interactive tool that illustrates how HHI relates to market share distributions.

## By Powertrain

### HHI by Make
```{r}
make_boxplot(data, "powertrain", "make")
```

### HHI by Vehicle Type
```{r}
make_boxplot(data, "powertrain", "vehicle_type")
```

### HHI by Price Bin
```{r}
make_boxplot(data, "powertrain", "price_bin")
```

## By Vehicle Type

### HHI by Make
```{r}
make_boxplot(data, "vehicle_type", "make")
```

### HHI by Price Bin
```{r}
make_boxplot(data, "vehicle_type", "price_bin")
```

## By Price Bin

### HHI by Make
```{r}
make_boxplot(data, "price_bin", "make")
```

### HHI by Vehicle Type
```{r}
make_boxplot(data, "price_bin", "vehicle_type")
```

## Data

<center>
[Download data](`r url_hhi`)

<div style="width:900px">

```{r}
data |>
  mutate(
    mean = round(mean, 3),
    median = round(median, 3),
    q25 = round(q25, 3),
    q75 = round(q75, 3),
    IQR = round(IQR, 3),
    upper = round(upper, 3),
    lower = round(lower, 3)
  ) |>
  rename(
    `Group Variable` = group_var,
    `Group Level` = group_level,
    `HHI Variable` = hhi_var,
    Year = listing_year,
    Mean = mean,
    Median = median,
    Q25 = q25,
    Q75 = q75,
    Upper = upper,
    Lower = lower
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50),
    defaultColDef = reactable::colDef(
      minWidth = 70
    )
  )
```

</div>
</center>
