---
title: "Percent of Dealers"
---

This page shows the percentage of dealerships that have at least one vehicle of a given type in their inventory over time. Each chart displays the share of dealers carrying a specific combination of powertrain and category (vehicle type or price bin), with separate views for new and used inventory. The x-axis is the listing year and the y-axis is the percent of dealers.

```{r}
#| include: false

source("setup.R")

# Load data - use URL since package may not have it yet
percent_dealers <- read.csv(url_percent_dealers)

data <- percent_dealers |>
  mutate(percent = round(p * 100, 2))

# Helper function to create a chart for a given group_level
make_chart <- function(data, grp_var, grp_level, cat_var, inv_type) {
  plot_data <- data |>
    filter(
      group_var == grp_var,
      group_level == grp_level,
      category_var == cat_var,
      inventory_type == inv_type
    )

  if (nrow(plot_data) == 0) {
    return(NULL)
  }

  plot_data |>
    mutate(listing_year = as.character(listing_year)) |>
    arrange(listing_year) |>
    group_by(category_level) |>
    e_charts(listing_year) |>
    e_line(percent, symbol = "circle", symbolSize = 6) |>
    e_tooltip(
      trigger = "axis",
      formatter = htmlwidgets::JS(
        "
        function(params) {
          var result = params[0].axisValueLabel + '<br/>';
          params.forEach(function(item) {
            result += item.marker + ' ' + item.seriesName + ': ' +
                      item.value[1].toFixed(2) + '%<br/>';
          });
          return result;
        }
      "
      )
    ) |>
    e_x_axis(
      name = "Listing Year",
      type = "category"
    ) |>
    e_y_axis(name = "% of Dealers") |>
    e_legend(
      show = TRUE,
      orient = "horizontal",
      left = "center",
      bottom = 0,
      width = "70%",
      selector = list(
        list(type = "all", title = "All"),
        list(type = "inverse", title = "Inverse")
      ),
      selectorPosition = "start"
    ) |>
    e_grid(bottom = 100) |>
    e_color(background = "transparent") |>
    e_toolbox_feature(feature = "dataZoom") |>
    e_toolbox_feature(feature = "saveAsImage")
}

# Helper to render a grid of charts for a group×category combo
render_combo <- function(data, grp_var, cat_var, inv_type) {
  levels <- data |>
    filter(
      group_var == grp_var,
      category_var == cat_var,
      inventory_type == inv_type
    ) |>
    distinct(group_level) |>
    pull(group_level) |>
    sort()

  charts <- lapply(levels, function(lvl) {
    chart <- make_chart(data, grp_var, lvl, cat_var, inv_type)
    if (is.null(chart)) {
      return(NULL)
    }
    htmltools::div(
      class = "g-col-6",
      htmltools::h4(lvl),
      chart
    )
  })

  # Remove NULLs
  charts <- Filter(Negate(is.null), charts)

  htmltools::div(class = "grid", charts)
}
```

## Powertrain by Vehicle Type

*Click legend items to show/hide lines.*

::: {.panel-tabset}

### New

```{r}
render_combo(data, "powertrain", "vehicle_type", "New")
```

### Used

```{r}
render_combo(data, "powertrain", "vehicle_type", "Used")
```

:::

## Powertrain by Price Bin

*Click legend items to show/hide lines.*

::: {.panel-tabset}

### New

```{r}
render_combo(data, "powertrain", "price_bin", "New")
```

### Used

```{r}
render_combo(data, "powertrain", "price_bin", "Used")
```

:::

## Vehicle Type by Price Bin

*Click legend items to show/hide lines.*

::: {.panel-tabset}

### New

```{r}
render_combo(data, "vehicle_type", "price_bin", "New")
```

### Used

```{r}
render_combo(data, "vehicle_type", "price_bin", "Used")
```

:::

## Data

<center>
[Download data](`r url_percent_dealers`)

<div style="width:900px">

```{r}
percent_dealers |>
  mutate(
    p = round(p, 4)
  ) |>
  rename(
    `Inventory Type` = inventory_type,
    `Group Variable` = group_var,
    `Group Level` = group_level,
    `Category Variable` = category_var,
    `Category Level` = category_level,
    `Listing Year` = listing_year,
    `Percent` = p
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50)
  )
```

</div>
</center>
