---
title: "Price Trends"
format:
  html:
    include-in-header:
      text: |
        <script src="https://cdn.jsdelivr.net/npm/echarts@5.4.3/dist/echarts.min.js"></script>
---

This page shows average vehicle price trends over time by state, vehicle type, and vehicle year. Select a state from the dropdown to explore pricing patterns for both new and used vehicles. New vehicle prices are shown for vehicles listed between 2018-2024, while used vehicle prices include vehicles from 2004-2024 with depreciation rate information.

```{r}
#| include: false

source("setup.R")

# Load price data - no filtering
data_new <- price_trends_new
data_used <- price_trends_used

# Get unique states and listing years (sorted)
states <- sort(unique(c(data_new$state, data_used$state)))
listing_years <- sort(unique(c(data_new$listing_year, data_used$listing_year)), decreasing = TRUE)
```

<div style="margin-bottom: 20px;">
  <label for="stateSelector" style="font-weight: bold; margin-right: 10px;">State:</label>
  <select id="stateSelector" style="padding: 5px 10px; font-size: 14px; border: 1px solid #ccc; border-radius: 4px;">
    <option value="CA">California</option>
  </select>

  <label for="listingYearSelector" style="font-weight: bold; margin-left: 20px; margin-right: 10px;">Listing Year:</label>
  <select id="listingYearSelector" style="padding: 5px 10px; font-size: 14px; border: 1px solid #ccc; border-radius: 4px;">
    <option value="2025">2025</option>
  </select>
</div>

::: {.panel-tabset}

## New Vehicles

<div style="margin-bottom: 20px;">
  <p><em>Average prices for new vehicles by vehicle type and model year. Data shown for listings from 2018-2024.</em></p>
</div>

```{r}
#| echo: false

library(jsonlite)

# Prepare data for JavaScript
data_new_json <- data_new |>
  mutate(
    listing_year = as.integer(listing_year),
    year = as.integer(year),
    avg_price = round(avg_price, 2)
  ) |>
  select(listing_year, state, year, vehicle_type, avg_price, n_listings) |>
  toJSON(dataframe = "rows")

# Create the chart container
htmltools::tags$div(
  id = "chart-new",
  style = "width: 100%; height: 600px;"
)
```

## Used Vehicles

<div style="margin-bottom: 20px;">
  <p><em>Average prices for used vehicles by vehicle type and model year. Data shown for listings from 2018-2024 and includes vehicle years from 2004-2024.</em></p>
</div>

```{r}
#| echo: false

# Prepare data for JavaScript
data_used_json <- data_used |>
  mutate(
    listing_year = as.integer(listing_year),
    year = as.integer(year),
    avg_price = round(avg_price, 2),
    dep_annual_rate = round(dep_annual_rate, 4)
  ) |>
  select(
    listing_year,
    state,
    year,
    vehicle_type,
    avg_price,
    n_listings,
    dep_annual_rate
  ) |>
  toJSON(dataframe = "rows")

# Create the chart container
htmltools::tags$div(
  id = "chart-used",
  style = "width: 100%; height: 600px;"
)
```

## Data

<center>
<div style="width:900px">

### New Vehicle Prices

```{r}
data_new |>
  mutate(avg_price = round(avg_price)) |>
  rename(
    "Listing Year" = listing_year,
    "State" = state,
    "Vehicle Year" = year,
    "Vehicle Type" = vehicle_type,
    "Average Price" = avg_price,
    "Number of Listings" = n_listings,
    "Number of Make/Model/Year" = n_make_model_year
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50)
  )
```

### Used Vehicle Prices

```{r}
data_used |>
  mutate(
    avg_price = round(avg_price),
    dep_annual_rate = round(dep_annual_rate, 3)
  ) |>
  rename(
    "Listing Year" = listing_year,
    "State" = state,
    "Vehicle Year" = year,
    "Vehicle Type" = vehicle_type,
    "Average Price" = avg_price,
    "Number of Listings" = n_listings,
    "Annual Depreciation Rate" = dep_annual_rate,
    "Number of Make/Model/Year" = n_make_model_year
  ) |>
  reactable::reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 25, 50)
  )
```

</div>
</center>

:::

```{r}
#| echo: false

# Inject the data and JavaScript
htmltools::tags$script(htmltools::HTML(sprintf(
  '
// Data
const dataNew = %s;
const dataUsed = %s;
const states = %s;
const listingYears = %s;

// Wait for DOM to be ready
document.addEventListener("DOMContentLoaded", function() {
  console.log("Data loaded - New:", dataNew.length, "Used:", dataUsed.length);

  // Populate state selector
  const stateSelector = document.getElementById("stateSelector");
  stateSelector.innerHTML = "";
  states.forEach(state => {
    const option = document.createElement("option");
    option.value = state;
    option.textContent = state;
    stateSelector.appendChild(option);
  });
  stateSelector.value = "CA";

  // Populate listing year selector
  const yearSelector = document.getElementById("listingYearSelector");
  yearSelector.innerHTML = "";
  listingYears.forEach(year => {
    const option = document.createElement("option");
    option.value = year;
    option.textContent = year;
    yearSelector.appendChild(option);
  });
  yearSelector.value = listingYears[0]; // Most recent year

  // Function to create chart
  function createChart(containerId, data, state, listingYear, isUsed = false) {
    const container = document.getElementById(containerId);
    if (!container) {
      console.error("Container not found:", containerId);
      return null;
    }

    // Filter data for selected state and listing year
    const filtered = data.filter(d => d.state === state && d.listing_year === parseInt(listingYear));

    if (filtered.length === 0) {
      console.warn("No data for state:", state);
      return null;
    }

    // Get unique vehicle types
    const vehicleTypes = [...new Set(filtered.map(d => d.vehicle_type))];

    // Prepare series data
    const series = vehicleTypes.map(vt => {
      const vtData = filtered
        .filter(d => d.vehicle_type === vt)
        .sort((a, b) => a.year - b.year);

      return {
        name: vt,
        type: "line",
        data: vtData.map(d => [d.year, d.avg_price]),
        symbol: "circle",
        symbolSize: 6,
        emphasis: { focus: "series" }
      };
    });

    // Get unique years for x-axis
    const years = [...new Set(filtered.map(d => d.year))].sort();

    // Create chart
    const chart = echarts.init(container);

    const option = {
      tooltip: {
        trigger: "axis",
        formatter: function(params) {
          let result = "Year: " + params[0].axisValue + "<br/>";
          params.forEach(item => {
            result += item.marker + " " + item.seriesName + ": $" +
                     item.value[1].toLocaleString() + "<br/>";
          });
          return result;
        }
      },
      legend: {
        data: vehicleTypes,
        orient: "horizontal",
        left: "center",
        bottom: 0,
        type: "scroll"
      },
      grid: {
        left: "10%%",
        right: "5%%",
        bottom: "80",
        top: "40",
        containLabel: true
      },
      toolbox: {
        feature: {
          dataZoom: { yAxisIndex: "none" },
          saveAsImage: {}
        }
      },
      xAxis: {
        type: "value",
        name: "Vehicle Year",
        nameLocation: "middle",
        nameGap: 30,
        min: Math.min(...years),
        max: Math.max(...years),
        axisLabel: {
          formatter: function(value) {
            return Math.round(value);
          }
        }
      },
      yAxis: {
        type: "value",
        name: "Average Price ($)",
        nameLocation: "middle",
        nameGap: 50,
        axisLabel: {
          formatter: function(value) {
            return "$" + (value / 1000).toFixed(0) + "k";
          }
        }
      },
      series: series
    };

    chart.setOption(option);

    // Responsive resize
    const resizeHandler = function() { chart.resize(); };
    window.addEventListener("resize", resizeHandler);

    return chart;
  }

  // Function to update charts
  function updateCharts() {
    const state = stateSelector.value;
    const listingYear = yearSelector.value;

    // Dispose old charts
    if (window.chartNew) window.chartNew.dispose();
    if (window.chartUsed) window.chartUsed.dispose();

    // Create new charts
    window.chartNew = createChart("chart-new", dataNew, state, listingYear, false);
    window.chartUsed = createChart("chart-used", dataUsed, state, listingYear, true);
  }

  // Initialize charts
  updateCharts();

  // Update charts when selectors change
  stateSelector.addEventListener("change", updateCharts);
  yearSelector.addEventListener("change", updateCharts);
});
',
  data_new_json,
  data_used_json,
  toJSON(states),
  toJSON(listing_years)
)))
```
